annals_manuscript

### 2020-04-15 annals_manuscript.Rmd
updated

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)
library(tidyverse)
library(cowplot)
library(vegan)
library(broom)
library(tsnemicrobiota)
library(latex2exp)
library(wesanderson)
library(table1)
library(knitr)
library(kableExtra)

@@

#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006) # only uses the baseline samples. This is n=1006
atlas.pseq <- subset_samples(atlas.pseq, !is.na(mnd), !is.na(nationality) )

## Incidence of MND
```{r}

#THIS IS NO LONGER NEEDED.

# This block uses the raw data to examin  the incidence of MND to ensure the classificaiton is correct. It assumes that the data is in the correct place so the setwd at the top meeds to be accurate. 
# setwd('~/globalmicrobiome') 
# # For MND  nrows = 196 as thwew are other things after 196
# mnd.i.15_49 <- as.data.frame(read.csv("./rawdata/MND_I_15_49.csv", header=TRUE, nrows = 196))

# mnd.i.50_69  <- as.data.frame(read.csv("./rawdata/MND_I_50_69.csv", header=TRUE, nrows = 196))
# mnd.i.70 <- as.data.frame(read.csv("./rawdata/MND_I_70plus.csv", header=TRUE, nrows = 196))

# mnd.i.70.test <- mnd.i.70
# # Run function and change name
# mnd.i.15_49 <- mb.gather(mnd.i.15_49)
# colnames(mnd.i.15_49)[2] <- "15-49"

# # Run function and change name
# mnd.i.50_69 <- mb.gather(mnd.i.50_69)
# colnames(mnd.i.50_69)[2] <- "50-69"

# # Run function and change name
# mnd.i.70 <- mb.gather(mnd.i.70)
# colnames(mnd.i.70)[2] <- "70"

# # Join them together.   
# mnd.i <- list( mnd.i.15_49,mnd.i.50_69,mnd.i.70 ) %>% reduce(left_join, by = "Location")

# a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE","Scandinavia" ))

# mnd.i %>%
#   kable(col.names = c("","Incidence","CI","Incidence","CI","Incidence","CI")) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"),full_width = F,position = "center") %>%
# add_header_above(c("Location" = 1, "15-49" = 2, "50-69" = 2, "70+" = 2)) %>%
#   add_header_above(c(" ", "Age (years)" = 6)) %>%
#   pack_rows("High MND group", 5,6)%>%
#   pack_rows("Low MND group", 2, 2) %>%
#   pack_rows("Medium MND group",1,3,4)

# # Now do the same for the curated data set
# mnd.ic.15_49 <- mb.cgather(mnd.i.15_49)
# colnames(mnd.ic.15_49)[2] <- "15-49"

# # Run function and change name
# mnd.ic.50_69 <- mb.cgather(mnd.i.50_69)
# colnames(mnd.ic.50_69)[2] <- "50-69"

# # Run function and change name
# mnd.ic.70 <- mb.cgather(mnd.i.70)
# colnames(mnd.ic.70)[2] <- "70"

# # Join them together.

# mnd.ic <- list( mnd.ic.15_49,mnd.ic.50_69,mnd.ic.70 ) %>% reduce(left_join, by = "Location")

# mnd.ic <- dplyr::select(mnd.ic, "Location",  "Value.x", "Lower.bound.x", "Upper.bound.x", "Value.y", "Lower.bound.y", "Upper.bound.y", "Value", "Lower.bound", "Upper.bound" )

# # Prep the data for the table

# mnd.ic %>%
#   kable(col.names = c("","Incidence","Upper","Lower","Incidence","Upper","Lower","Incidence","Upper","Lower"),digits=3) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"),full_width = F,position = "left") %>%
# add_header_above(c("Location" = 1, "15-49" = 3, "50-69" = 3, "70+" = 3)) %>%
#   add_header_above(c(" ", "Age (years)" = 9)) %>%
#   pack_rows("Group 1", 4, 7) %>%
#   pack_rows("Group 1", 4, 7) %>%
#   pack_rows("Group 2", 8, 10)

@@

my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category | mnd, data=data)
}

my_atlas_table <- function(atlas.pseq){
    n <- length(sample_data(atlas.pseq)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(atlas.pseq)$mnd
    data$bmi <- sample_data(atlas.pseq)$bmi_group
    data$age <- sample_data(atlas.pseq)$age
    table1(~ age  +  bmi | mnd, data=data)
}

@@

# https://bioconductor.riken.jp/packages/3.5/data/experiment/vignettes/curatedMetagenomicData/inst/doc/curatedMetagenomicData.html
# Prepare data for vizualisation

@@

atlas.pseq <- subset_samples(atlas.pseq, !is.na(nationality), !is.na(mnd))

@@

```{r Abundance}

@@

p <- plot.p.new.rel + scale_fill_brewer(palette = "Paired")
ggsave("atlas_bar.png")

p1 <- plot.p.new.rel + scale_fill_manual(values=c( "#A5CEE3", "#1F78B4", "#34A02C", "#E3211D", "#FF7E00", "blue", "yellow", "red"))
ggsave("pseq_bar.png")

p3 <<-plot_grid(title,p,p1, align = "v" , ncol = 1, labels = c('','A', 'B'),rel_heights = c(0.1, 1,1.2))
plot(p3)

@@

# Alpha and beta diversity figure
```{r}

@@

# gets the meta data 
  pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

@@

  # This country versus nationality
  p3<-   ggviolin(pseq.meta, x = "nationality", y = "Shannon",
                  add = "boxplot", fill = "nationality", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")

@@

  p4 <<- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red"))

  p4 <<- p4 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND DALYs") + theme(legend.position="none")

p.1 <<-plot_grid(title, p1, p3, p2, p4, align = "v" , ncol = 2, labels = c('','A', 'B', 'C', 'D'), rel_widths = 1, rel_heights = 2)
print(p.1)

>>>>

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)
library(tidyverse)
library(cowplot)
library(vegan)
library(broom)
library(tsnemicrobiota)
library(latex2exp)
library(wesanderson)
library(table1)
library(knitr)
library(kableExtra)
library(ggplot2)
library(grid)
library(gridExtra)

@@

#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006) # only uses the baseline samples. This is n=1006

@@

my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
     data$country <- sample_data(pseq.subset)$country
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category + country | mnd, data=data)
}

my_atlas_table <- function(atlas.pseq){
    n <- length(sample_data(atlas.pseq)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(atlas.pseq)$mnd
    data$bmi <- sample_data(atlas.pseq)$bmi_group
     data$nationality <- sample_data(atlas.pseq)$nationality
    data$age <- sample_data(atlas.pseq)$age
    table1(~ age  +  bmi + nationality | mnd, data=data)
}

@@

#### FIkter curated
# https://bioconductor.riken.jp/packages/3.5/data/experiment/vignettes/curatedMetagenomicData/inst/doc/curatedMetagenomicData.html
# Prepare data for vizualisation

@@

#atlas.pseq <- subset_samples(atlas.pseq, !is.na(nationality), !is.na(mnd))

#### Filter Atlas
atlas.pseq <- subset_samples(atlas.pseq, !is.na(mnd), !is.na(nationality))

#### List of genera in phylums
```{r List of genera in phylum}

map_levels(NULL, "Phylum", "Genus", pseq.subset)

map_levels(NULL, "Phylum", "Genus", atlas.pseq)

#taxa <- map_levels(NULL, "Phylum", "Genus", atlas.pseq)$Clostridium_cluster_I
#taxa <- map_levels(NULL, "Phylum", "Genus", atlas.pseq)$Actinobacteria
# List of Genera in the Bacteroideted Phylum
#taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
#print(taxa)

### Clostridium grouping NOT NEEDED ANYMORE since groupoings have been updated
```{r}

# Clostridum grouping

#clostridium <- subset_taxa(atlas.pseq, Phylum =="Clostridium cluster I" | Phylum =="Clostridium cluster XI"| Phylum =="Clostridium cluster XVI"  | Phylum =="Clostridium cluster III"  | Phylum =="Clostridium cluster XIII" | Phylum =="Clostridium cluster XVII"| Phylum =="Clostridium cluster IV" | Phylum =="Clostridium cluster XIVa" | Phylum =="Clostridium cluster XVIII" | Phylum =="Clostridium cluster IX" | Phylum =="Clostridium cluster XV" | Phylum =="Uncultured Clostridiales")

#atlas.pseq.new <- subset_taxa(atlas.pseq, Phylum =="Actinobacteria" | Phylum =="Bacilli"| Phylum =="Proteobacteria"  | Phylum =="Verrucomicrobia"  | Phylum =="Bacteroidetes" | Phylum =="Asteroleplasma"| Phylum =="Spirochaetes" | Phylum =="Fusobacteria" | Phylum =="Cyanobacteria" | Phylum =="Uncultured Mollicutes")

## Summary
```{r summary}
# Summarise contents of phyloseq
summarize_phyloseq(pseq)
# Sparsity is how populated is the data with zeros
kable(head(tax_table(pseq)))
tax_table(pseq)
# Check the data
print(pseq)
# Display variables from the mapping file
sample_variables(pseq) 
# Total number of taxa in the entire data
ntaxa(pseq) 
# Taxonomic ranks
rank_names(pseq) 
```

```{r}

@@

#Abundance function
```{r}

@@

@@

#Abundance plots figure
```{r Abundance}

plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = "MND DALYs", plot.type = "barplot", title = "Atlas Abundance Bar Plot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic")) + theme_classic()  + labs(x="MND") + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 10)) + scale_x_discrete(labels=c("LOW" = "Low", "MEDIUM" = "Medium",
                              "HIGH" = "High"))

###+ theme(legend.title = element_text(size = 7), legend.text = element_text(size = 7))

p <- plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
#ggsave("atlas_bar.png")

plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = "MND DALYs", plot.type = "barplot", title = "Curated Abundance Bar Plot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text(face="italic")) + theme_classic() + labs(x="MND") + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 10)) + theme(legend.position="none")  + scale_x_discrete(labels=c("LOW" = "Low", "MEDIUM" = "Medium","HIGH" = "High"))

p1 <- plot.p.new.rel + scale_fill_manual(values=c( "#A5CEE3", "#1F78B4", "#34A02C", "#E3211D", "#FF7E00", "blue", "yellow", "red"))
#ggsave("pseq_bar.png") + scale_fill_brewer(palette = "Paired") + labs(title = "B")

# 2. Save the legend
#+++++++++++++++++++++++
legend <- get_legend(p)
# 3. Remove the legend from the box plot
#+++++++++++++++++++++++
p<- p + theme(legend.position="none")
# 4. Arrange ggplot2 graphs with a specific width
abun_plot <-grid.arrange(p, p1, legend, ncol=3, widths=c(2.5, 2.5, 1))

#abun_plot <- plot_grid(p,p1, align = "h", axis = "bt", ncol = 2, nrow=1, labels = c('A', 'B'), rel_heights = c(10,10)) 
abun_plot
plot(abun_plot)
ggsave(file = "bar_plots.png", dpi = 600, width = 10, height = 5, units = "in")

ggsave("bar_plots.png", width = 10, height = 5)

#,rel_heights = c(30, 30, 30)

#t3 <- grid_arrange_shared_legend(p, p1 , ncol = 2, nrow = 1, position='right')
#t3

#g_legend<-function(a.gplot){
#  tmp <- ggplot_gtable(ggplot_build(a.gplot))
#  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
#  legend <- tmp$grobs[[leg]]
#  return(legend)}

#library(gridExtra)
#legend <- g_legend(p + theme(legend.position='right'))
#f <- grid.arrange(p+theme(legend.position='hidden'), p1+theme(legend.position='hidden'), ncol = 2, nrow=1, legend)

# Alpha figure of nationality
```{r}

  # creates Shannon
  pseq.meta$Shannon <- tab$diversity_shannon

  # This country versus nationality
  p <- ggviolin(pseq.meta, x = "nationality", y = "Shannon", add = "boxplot", fill = "nationality", palette = "paired", title = "Atlas Nationality Violin Plot") + theme(axis.text.x = element_text(angle = 90)) + xlab("") +theme(legend.direction = "vertical") + theme( axis.text.x=element_blank()) + labs(fill = "Nationality") + xlab("Nationality") + theme(title = element_text(size = 8))

pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  # creates Shannon
  pseq.meta$Shannon <- tab$diversity_shannon

  # This country versus nationality
  p1 <- ggviolin(pseq.meta, x = "nationality", y = "Shannon",
                  add = "boxplot", fill = "nationality", palette = "paired", title = "Curated Nationality Violin Plot") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("") + theme( axis.text.x=element_blank()) + labs(fill = "Nationality") + xlab("Nationality") + theme(title = element_text(size = 8))

legend <- get_legend(p) 
# 3. Remove the legend from the box plot
#+++++++++++++++++++++++
p<- p + theme(legend.position="none")
# 4. Arrange ggplot2 graphs with a specific width
p2 <- grid.arrange(p, p1, legend, ncol=3, widths=c(10, 10, 3))

#lay <- rbind(c(p, p1),
            # c(legend))

p2 <- grid.arrange(p, p1, legend, ncol=2, nrow = 2, layout_matrix = rbind(c(1,2), c(3,3)), widths = c(2.7, 2.7), heights = c(2.5, 0.2), legend.direction = "vertical")

ggsave("nationality_violin.png", width = 10, height = 5)

# Alpha figure of nationality and MND
```{r}

pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  print("Second graph has completed")

  p3 <<-plot_grid(p1,p2, align = "v" , ncol = 1, axis = 'tblr', labels = c('D', 'F'),rel_heights = c(10,10))

print(p3)

@@

p4 <<-plot_grid(p1,p2, align = "v" , ncol = 1, axis = 'tblr', labels = c('C', 'E'),rel_heights = c(10,10))

print(p4)

#order is atlas first

p6 <<-plot_grid(p4,p3, align = "v" , ncol = 2, rel_heights = c(20,20))

print(p6)

ggsave("alpha_manuscript.png", width = 10, height = 10)

# Composition and Alpha 
```{r}

p8 <<-plot_grid(abun_plot, p6, align = "v" , nrow=2, rel_heights = c(12,30,30))

print(p8)

ggsave("alpha_comp_manuscript.png", width = 10, height = 15)

theme(axis.title.x=element_blank(), axis.text.x=element_blank()
# Alpha figure of just MND
```{r}

pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  print("The basic alpha diversity analysis has been completed")

  p <<- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red")) + theme( axis.text.x=element_blank()) + scale_fill_discrete(name = "MND DALYs", labels = c("Low", "Medium", "High"))

  p <<- p + stat_compare_means(comparisons = mnd.pairs) + xlab("MND DALYs") + theme(legend.position="none")

  print("Second graph has completed")

@@

@@

  p2 <<- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red")) + theme( axis.text.x=element_blank()) + labs(fill = "MND") + scale_fill_discrete(name = "MND DALYs", labels = c("Low", "Medium", "High"))

  p2 <<- p2 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND DALYs") + theme(legend.position="right")

# 2. Save the legend
#+++++++++++++++++++++++
legend <- get_legend(p2)
# 3. Remove the legend from the box plot
#+++++++++++++++++++++++
p2<- p2 + theme(legend.position="none")
# 4. Arrange ggplot2 graphs with a specific width
p3 <- grid.arrange(p2, p, legend, ncol=3, widths=c(2.3, 2.3, 0.8))

#If want legend bottom centre 
# grid.arrange(p2, p, legend, ncol=2, nrow = 2, layout_matrix = rbind(c(1,2), c(3,3)), widths = c(2.7, 2.7), heights = c(2.5, 0.2), legend.direction = "horizontal")

  #p3 <<-plot_grid(p2,p, align = "v" , ncol = 2, axis = 'tblr', labels = c('C', 'D'),rel_heights = c(10,10))

print(p3)

ggsave("alpha2_manuscript.png", width = 10, height = 10)

# Composition and Alpha  of just MND
```{r}

p8 <<-plot_grid(abun_plot, p3, align = "v" , nrow=2, rel_heights = c(30,30))

print(p8)

ggsave("alpha_comp_manuscript2.png", width = 10, height = 10)

#Beta
```{r}

 p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

 p <<-  boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

  # This add the facets but I do not think we need that any more. 
  #p3 <- p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
  ###save_plot(paste("Beta_diversity_",title_text,"_NMDS.png", sep = "" ), p3)
    # MDS
   pseq.mds <<- ordinate(pseq.subset, "MDS", "bray")
   p4 <- plot_ordination(pseq.subset, pseq.mds, type = "sample", color = "mnd")
 #plot(p1 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence MDS"))

    p5 <- p4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND DALYs") + theme(legend.position = "none")
    ###save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p5)

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND DALYs") + theme(legend.position = "bottom")
   ### save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p8)

    p10 <<-plot_grid(title,p3, p5,p8, ncol = 1, labels = c('','A', 'B','C'),rel_heights = c(0.3, 1.5,1.5,2))

    print(p10)

# atlas
 p.low <<- subset_samples(atlas.pseq, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(atlas.pseq, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(atlas.pseq, mnd == "HIGH")
  p.high.d <<- divergence(p.high)

 t <<-  boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
  pseq.nmds <- ordinate(atlas.pseq, "NMDS", "bray")
  t2 <- plot_ordination(atlas.pseq, pseq.nmds, type = "sample", color = "mnd")

  t3 <- pt + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "NMDS",color = "MND DALYs") + theme(legend.position = "none")
  #plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))

  # This add the facets but I do not think we need that any more. 
  #p3 <- p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")
  save_plot(paste("Beta_diversity_",title_text,"_NMDS.png", sep = "" ), p3)
    # MDS
   pseq.mds <<- ordinate(atlas.pseq, "MDS", "bray")
   t4 <- plot_ordination(atlas.pseq, pseq.mds, type = "sample", color = "mnd")
 #plot(p1 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence MDS"))

    t5 <- t4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND DALYs") + theme(legend.position = "none")
   ### save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), t5)

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(atlas.pseq, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    t7 <- plot_tsne_phyloseq(atlas.pseq, tsne_res,color = 'mnd')
    t8 <- t7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND DALYs") + theme(legend.position = "bottom")
### save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), t8)

    t10 <<-plot_grid(title,t3, t5, t8, ncol = 1, labels = c('','A', 'B','C'),rel_heights = c(0.3, 1.5,1.5,2))

    print(t10)

beta <<-plot_grid(p10, t10, align = "h" , ncolumn = 2, rel_heights = c(30,30,30))

beta

# Permanova figure
```{r}

 p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = TRUE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Potenital Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme_classic() + theme(axis.text.y = element_text(face="italic")) + ggtitle("Curated PERMANOVA")

plot(p, xlim=c(-0.03,0.03))

  # scale_x_continuous(limits=c(-0.03, 0.03), breaks=c(-0.03, -0.02, -0.01, 0, 0.01, 0.02, 0.03))

 p.low <<- subset_samples(atlas.pseq, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(atlas.pseq, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(atlas.pseq, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p2 <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = TRUE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Potential Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme_classic() + theme(axis.text.y = element_text(face="italic")) + ggtitle("Atlas PERMANOVA")

perm <- plot_grid(p2,p, align = "v" , ncol = 1, axis = 'tblr', labels = c('A', 'B'),rel_heights = c(10,10,10))

print(perm)
ggsave("permanova_annals.png", device = "png",width = 250, height = 280, units = "mm")
